\mytitle

### 2020-04-18 paper.tex
pass over abstract, intro, 2,3.

\begin{abstract} 
%Wind farms play an increasingly important
%role as a source of renewable energy and 
The wind farm layout optimization (WFLO) problem requires the
selection of wind turbine positions to maximize energy production. While a number of heuristic and exact techniques have been proposed in the literature to solve WFLO, it is natural to model the interaction between turbines as a quadratic objective function as, indeed, has been done in some existing optimization models. Recent advances in optimization hardware and software have both targeted quadratic constraints: commercial mixed integer linear solvers have been extended to address some quadratic problems and nascent specialized hardware, including quantum computing systems, have focused on solving Quadratic Unconstrained Binary Optimization (QUBO) problems. 
%It is a key problem 
%when designing a new wind farm, as it has a direct influence
%on the amount of the total produced energy. 
%In recent years there have been advances in both software technologies (e.g., the extension of commercial mixed integer linear programming solvers to handle quadratic problems) and nascent hardware development (e.g., quantum computing chips and specialized classical chips) that have relevance for the WFLO problem. In both cases, the advances have 
%Recent years saw 
%advances in software and hardware optimization technologies, such as Gurobi's quadratic solver 
%and Fujitsu's digital annealer (DA), respectively. 
%These technologies create an opportunity to improve 
%over current wind farm optimization solutions
%both in terms of solution quality and solve time. 
%In order to exploit these technologies, 
%we must represent the WFLO problem as 
%a quadratic optimization model. Specifically, in order 
%to apply a software-based optimization
%solution (Gurobi), 
In this paper, we 
introduce two novel quadratic programming models for WFLO: a quadratic constrained optimization problem (QCOP) with binary decision variables and a QUBO.
%Subsequently, we provide the unconstrained counterpart of the QCOP, 
%namely the quadratic unconstrained binary optimization (QUBO), which  
%is the underlying representation for quantum-inspired 
%optimization hardware such as the DA. 
A thorough empirical 
evaluation using a commercial solver and a specialized classical chip for QUBO solving demonstrates the strengths of the two models. 
Our results show that both approaches yield fast and high-quality solutions 
and
improve over existing WFLO solutions. In particular, the QUBO model delivers high quality solutions very quickly while the QCOP model can be used to find better solutions and provide quality guarantees over a longer run-time.
\end{abstract}
\begin{keyword}
	Wind Farm Layout Optimization (WFLO) \sep Quadratic Programming (QP) \sep Quadratic Unconstrained Binary Optimization (QUBO) \sep Digital Annealer
\end{keyword}

@@

\section{Introduction}
% Todo
The goal in the wind farm layout optimization (WFLO)
problem~\cite{MOSETTI1994105} is to locate wind turbines within a
predefined area to maximize the total power harvested from the wind
stream.  While physical considerations based on the diameter of the
blades prevent turbines being placed too close to each other, turbine
location decisions influence the total power production due to
interference effects (i.e., wakes) generated by upstream turbines.
These inter-turbine effects are naturally captured using quadratic
constraints and objective functions. However, with a few exceptions (e.g., \cite{turner2014new}), exact optimization
approaches that have been proposed for the WFLO
problem~\cite{Zhang2014,kuo2016wind} \todo{[JCB: Add more references here]}
represent the quadratic relations using linear models
that approximate the true dependencies.

In this work, we introduce two novel quadratic WFLO models. The first
model represents WFLO with a quadratic constrained optimization
problem (QCOP).  The QCOP can be solved using state-of-the-art solvers
that specialize in such problems, e.g., Gurobi~\cite{gurobi}.  Our
second model is based on a quadratic unconstrained binary optimization
(QUBO) representation of the WFLO problem.  Such a formulation enables the use of
specialized optimization hardware tailored to solve QUBOs. In this
paper, we use Fujitsu's Digital Annealer (DA), a hardware-based
QUBO solver, to demonstrate the strength of our approach. The main
contributions of this paper are:
\begin{enumerate} 
\item We propose a novel quadratic modeling framework for WFLO that captures wake 
effects directly through their linear superposition. 
%without the need for approximations and ad-hoc solutions,
\item We prove that WFLO is $\mathcal{NP}$-hard.
\item Through numerical experiments on two instantiations of our quadratic framework using Fujitsu's Digital Annealer and a commercial software-based quadratic solver (Gurobi), we demonstrate state-of-the-art performance in, respectively, quickly finding high-quality solutions and finding and proving optimal solutions over a longer run-time.
\end{enumerate}

@@

The remainder is structured as follows \todo{adjust structure according to final structure of paper}.

@@

%\paragraph{Wind Farm as a Grid}
We consider a square-shaped wind farm that we model as an $n\times n$
grid with $n$ being the number of cells on each axis.  The cells have
an area of $c^2$ (in squared meters), i.e., if $c=100$ meters and
$n=20$, the farm area will be $2km\times 2km$.  In this work, we
assume that the terrain is flat and that we have no constraints
related to the terrain.  Furthermore, we do not account for the
influence of noise on the wind farm's surroundings. The wind farm
literature treats the more complex farm models that capture non-flat
terrain~\cite{song2015lazy,kuo2016wind}, multi-typed
turbines~\cite{feng2017design}, noise
effects~\cite{Zhang2014,sorkhabi2016impact,yin2014multi}, and complex
objectives (including turbine installation and maintenance
costs)~\cite{lackner2007analytical}.  In this work, we focus on the
novel quadratic approach for WFLO, and hence consider the most
abstract wind farm model. Our model can be extended, in the
future, to include additional wind farm settings. \todo{perhaps
  discuss in future work}

@@

\paragraph{Effective Power Calculations in the Presence of Wakes} 
Let $\mathcal{D}$ be the set of wind regimes where every regime, $d\in \mathcal{D}$, is a unique \todo{[JCB: Mathematically, I don't think it matters if the regime is unique. So delete the word?]} combination of wind direction and free speed. 
The probability of wind regime 
$d \in \mathcal{D}$ is given by $p_d$ with $\sum_{d \in \mathcal{D}}^{} p_d = 1$. In the experimental section (Section~\ref{sec:eval}),
we provide one example of a probability distribution 
function over $36$ wind regimes (Figure~\ref{fig:prob_wind}). In that example, 
a regime $d=(10^\circ, 12 \mbox{ km/h})$ corresponds to a north-eastern wind blowing at a free velocity of $12\mbox{ km/h}$. The
probability of this regime is $0.008$.

@@

  Eq. (\ref{eq:ess}) is considered the most accurate analytical total power expression that accounts for multiple wakes.\todo{[JCB: REF?]} However, the expression is not amenable to optimization \todo{[JCB: because?]}. Hence, a simplified representation for the total expected energy that considers wakes is required.

@@

Note that $E_{LS}$ has the same parameters as $E_{SS}$ but differs in their relationship.
The LS model is less accurate compared to SS,\todo{[JCB: REF?]} yet as we will show in Section~\ref{sec:QUBO4WFLO} it 
results in a quadratic objective function in our WFLO formulations, which is a desired property from an optimization standpoint. \todo{[JCB: A bit strange to claim that a quadratic model is ``natural'' in the abstract but then argue that the model we adopt is an approximation but it is good because it is quadratic. Sounds like it is not natural and we choose it due to its optimization properties.]}

@@

In what follows, we provide a literature review that covers both exact and approximate methods for WFLO. We subsequently relate our 
work to both approaches via our proposed quadratic optimization framework for WFLO.\footnote{We review 
only literature that considers the same wind farm model as the one presented at the beginning of this section. For a broader literature review see \todo{[REF]}.}

\paragraph{Exact Optimization Methods} 
\todo{Rephrase and add citations to exact methods}
The WFLO problem can be solved  
to optimality, using
techniques from operations research (e.g., mixed-integer linear programming (MILP), quadratic programming (QP), and constraint programming (CP))~\cite{}.
These methods are \emph{exact}, in the sense that they guarantee that the returned solution is indeed
the best solution one can obtain for the given problem. Donovan was the first to solve WFLO using a
MILP approach based on the LS energy expression~\cite{donovan2005wind}. In~\cite{turner2014new}, the authors
approximated the SS energy expression using a quadratic and a linear approximation. The two resulting methods performed well
compared to existing approximate and exact methods~\cite{turner2014new}.

The most recent attempt to solve the wind farm variation that we presented 
in this section is given in~\cite{Zhang2014}. The authors compared
previously existing MILP and QP models to their approach for directly optimizing the SS objective. To this end, they used constraint programming (CP), which 
is an optimization technique that does not make assumptions on the functional form of the objective.
This turned out to be computationally intractable for the larger standard WFLO benchmarks.

In our work, we build upon the approach in~\cite{donovan2005wind}, yet instead of using a linear version of the LS objective, we 
model WFLO using a more natural quadratic representation. The difference between our approach and the quadratic model
proposed in~\cite{turner2014new} is that our model is not an approximation, but rather an exact representation of the LS objective.

@@

\paragraph{Approximate Methods} \todo{tighten} In contrast to exact methods, approximation algorithms often provide   
quick and sufficiently accurate solutions. Several works solved 
WFLO using evolutionary (genetic) approaches~\cite{MOSETTI1994105,gonzalez2010optimization,grady2005placement}. 
These methods use two principles (cross-over and mutation)
to change existing solutions and improve them over time. Evolutionary methods do not guarantee termination 
(i.e., they may run forever without achieving a quality criteria), and when they do terminate (in practical cases) they do
not guarantee optimal solutions. Hence, one must often
introduce non-quality related stopping criteria, 
thus compromising on quality of the attained solution~\cite{davis1991handbook}.
Furthermore, there is no natural way of introducing constraints into genetic methods, which
leads to a potentially costly feasibility check for every solution.

The first attempt to address these limitations 
was made in a local search approach for WFLO~\cite{ozturk2004heuristic}. 
The solution does not guarantee optimality, 
yet it circumvents the termination and feasibility checking issues of the evolutionary
methods. The local search approach attempts to apply either move, remove or add actions given an incumbent 
feasible solution. When a turbine is moved, removed or added, the new solution is evaluated. The procedure terminates 
when a stopping criteria is reached.
The main limitation of~\cite{ozturk2004heuristic} is that it cannot perform an action 
that leads to an infeasible state. Therefore, it quickly converges to a sub-optimal solution (local optimum) 
that can potentially by far
from the globally optimal solution. This problem was experimentally demonstrated in~\cite{rivas2009solving}.

To overcome this limitation, the authors of~\cite{rivas2009solving} used simulated annealing (SA), 
a neighborhood search method that allows visits to infeasible solutions. 
This creates a well-connected solution space that leads the algorithm to escape 
local optima. The SA method was shown to be superior to~\cite{ozturk2004heuristic} 
and to the genetic method proposed in~\cite{grady2005placement}. A drawback of the SA approach 
is its ad-hoc nature. The various components of the algorithm
 must be tailored and geared towards WFLO. This means that small adjustments to our WFLO setting
 (e.g., adding noise considerations), would require major changes to the SA solution. 
Additional methods and experimental comparisons between various approximate algorithms are available in~\cite{samorani2013wind}.

In our work, we also use an SA approach (similarly to~\cite{rivas2009solving}). However,
our methodology is generic since it is based on our quadratic paradigm and
on the ability of specialized hardware to solve quadratic (non-ad-hoc) models.
Quadratic WFLO models can be easily adjusted to include additional problem features, e.g., wind farm characteristics.

\paragraph{Quadratic Programming as a Unified Framework for Exact and Approximate WFLO}
\todo{send a clear message here. Perhaps move to next section}.

In this work, we argue that 
formulating WFLO using quadratic programming enables us to employ 
advanced software solutions (e.g., Gurobi optimizer) and approximate hardware developments 
(e.g., Fujitsu's digital annealer that solves quadratic problems using principles from simulated annealing). 
A benchmark study compared state of the art exact approaches from~\cite{Zhang2014} to 
approximate methods and found that the two perform in a comparable fashion, without large differences between
the attained energy~\cite{yang2019simulated}. Exact methods tend to perform slightly better
in smaller WFLO benchmarks, while simulated annealing methods dominated for larger standard WFLO benchmarks. 
This leads to the conclusion that both types of methods (exact and approximate) should be considered. 
Therefore, in this work, we use a unified paradigm that would enable both exact and approximate solutions to WFLO. Thus, 
our approach `enjoys' the best of both worlds.

@@

In this part, 
we formulate the WFLO problem
using quadratic programming. Specifically, we consider two
quadratic optimization problems that represent WFLO: a
quadratic constrained optimization (QCOP) problem and 
a quadratic unconstrained binary optimization (QUBO) problem. 
The former represents WFLO characteristics (proximity and number of turbines)
as hard constraints, while the latter considers these characteristics as soft constraints that are added to its objective function.

@@

Let $u_{id}$ and $u_{id\infty}$ 
be the wind speeds at turbine location $i \in \mathcal{N}$ for wind regime $d \in \mathcal{D}$
with and without interference from other turbines due to wake effects, respectively. 
Note that here we refer to $i$ as a potential turbine location and not the $i$th turbine as we did in Section~\ref{sec:related}.
Further, we denote by $\mathcal{U}_{id} \subseteq \mathcal{N}$ 
the set of upstream 
turbine locations for wind regime $d$
given that a turbine is placed at cell $i$ (i.e., turbines placed in these locations 
will be influenced by a turbine placed in $i$). Finally,
we denote by $u_{ijd}$ the wind speed at location $j$ due to a single wake from upstream turbine at location $i$ with $j \in \mathcal{U}_{id}$. 
The following function  
represents the total expected energy of a wind farm given placement decisions 
$x$: \begin{equation}
f(x) = \sum_{i \in \mathcal{N}}^{} \sum_{d \in \mathcal{D}}^{} p_d ( \frac{1}{3} \ u_{id, \infty}^3 x_i  - \sum_{j \in \mathcal{U}_{id}}^{} \frac{1}{3}(u_{id, \infty} ^3 - u_{ijd}^3)x_i x_j).   \label{ObjFunc}\\
%&s.t.:& \nonumber\\
%&\mbox{       }& \sum_{i \in \mathcal{N}}^{} x_i = m,\label{Cardinality}\\
%&\mbox{       }& x_i + x_j \leq 1,   \forall (i,j) \in \mathcal{E}, \label{Grid}\\
%&\mbox{       }& x_i \in \{0,1\},     \forall i \in \mathcal{N}.
\end{equation} The objective function corresponds to the linear superposition (LS)
expression (Eq.~\ref{eq:ls}). We are now ready to provide the two quadratic programs to represent WFLO.

@@

To 
maximize $f(x)$ and satisfy the total number of turbines and proximity 
constraints we write
the following quadratic optimization problem (QCOP):
\begin{eqnarray} \label{QCOP}
&\max_{x_i}^{}& \sum_{i \in \mathcal{N}}^{} \sum_{d \in \mathcal{D}}^{} p_d ( \frac{1}{3} \ u_{id, \infty}^3 x_i  - \sum_{j \in \mathcal{U}_{id}}^{} \frac{1}{3}(u_{id, \infty} ^3 - u_{ijd}^3)x_i x_j)   \\
&s.t.:& \nonumber\\
&\mbox{       }& \sum_{i \in \mathcal{N}}^{} x_i = m,\label{Cardinality}\\
&\mbox{       }& x_i + x_j \leq 1,   \forall (i,j) \in \mathcal{E}, \label{Grid}\\
&\mbox{       }& x_i \in \{0,1\},     \forall i \in \mathcal{N}.
\end{eqnarray} Constraint~\ref{Cardinality} and \ref{Grid} ensure that exactly $m$ turbines are placed on the grid
and enforce proximity constraints between the turbines, respectively.

The QCOP presented as Problem~\ref{QCOP} can be solved using an exact optimization solver, e.g., Gurobi~\cite{gurobi}. 
So far, in~\cite{Zhang2014} and~\cite{donovan2005wind}, equivalent (linear) problems were formulated and solved using 
exact solvers. However, these solvers should only be used if the problem is computationally hard.
We are unaware of an existing complexity result for the LS variation in Eq.~\ref{QCOP}. In what follows, we provide
a theorem that proves the computational complexity of WFLO. 
\begin{mythm}
	The computational complexity of WFLO is $\mathcal{NP}$-hard. 
\end{mythm}
\begin{myproof}
We prove by showing that a special case of the WFLO problem (Eq.~(\ref{QCOP}) is
$\mathcal{NP}$-hard. Suppose that $\mathcal{E} = \emptyset$, i.e., the grid is coarse grained enough to ignore proximity constraints. Then, the constraints in Eq.~(\ref{Grid})
are dropped and we get the heaviest k-subgraph problem (HSP), which was proven to be $\mathcal{NP}$-hard in~\cite{billionnet2005different}.
Since a special case of WFLO is  $\mathcal{NP}$-hard it implies that WFLO is at least $\mathcal{NP}$-hard. On the other hand,
WFLO is formulated using integer programming (Problem \ref{QCOP}), which means that it is at most $\mathcal{NP}$-hard. Therefore, we get that WFLO is $\mathcal{NP}$-hard.
	\end{myproof} To overcome the challenge associated with the computational hardness of WFLO,
we reformulate the QCOP into a QUBO problem.

@@

\vspace{0.5em}
\caption{Total Expected Power (kilo Watt/year) after 3600 seconds as function of benchmark and method}\label{tab:results1}
\end{table}

@@

>>>>

\begin{abstract} 
%Wind farms play an increasingly important
%role as a source of renewable energy and 
The wind farm layout optimization (WFLO) problem requires the
selection of wind turbine positions to maximize energy production. While a number of heuristic and exact techniques have been proposed in the literature to solve WFLO, it is common to model the interaction between turbines as a quadratic objective function as, indeed, has been done in some existing optimization models. Recent advances in optimization hardware and software have both targeted quadratic constraints: commercial mixed integer linear solvers have been extended to address some quadratic problems and nascent specialized hardware, including quantum computing systems, have focused on solving Quadratic Unconstrained Binary Optimization (QUBO) problems. 
%It is a key problem 
%when designing a new wind farm, as it has a direct influence
%on the amount of the total produced energy. 
%In recent years there have been advances in both software technologies (e.g., the extension of commercial mixed integer linear programming solvers to handle quadratic problems) and nascent hardware development (e.g., quantum computing chips and specialized classical chips) that have relevance for the WFLO problem. In both cases, the advances have 
%Recent years saw 
%advances in software and hardware optimization technologies, such as Gurobi's quadratic solver 
%and Fujitsu's digital annealer (DA), respectively. 
%These technologies create an opportunity to improve 
%over current wind farm optimization solutions
%both in terms of solution quality and solve time. 
%In order to exploit these technologies, 
%we must represent the WFLO problem as 
%a quadratic optimization model. Specifically, in order 
%to apply a software-based optimization
%solution (Gurobi), 
In this paper, we 
introduce two novel quadratic programming models for WFLO: a quadratic constrained optimization problem (QCOP) with binary decision variables and a QUBO.
%Subsequently, we provide the unconstrained counterpart of the QCOP, 
%namely the quadratic unconstrained binary optimization (QUBO), which  
%is the underlying representation for quantum-inspired 
%optimization hardware such as the DA. 
A thorough empirical 
evaluation using a commercial solver and a specialized classical chip for QUBO solving demonstrates the strengths of the two models. 
Our results show that both approaches yield fast and high-quality solutions 
and
improve over existing WFLO solutions. In particular, the QUBO model delivers high quality solutions very quickly while the QCOP model can be used to find better solutions and provide quality guarantees over a longer run-time.
\end{abstract}
\begin{keyword}
	Wind Farm Layout Optimization (WFLO) \sep Quadratic Programming (QP) \sep Quadratic Unconstrained Binary Optimization (QUBO) \sep Digital Annealer
\end{keyword}

@@

\section{Introduction}
% Todo
The goal in the wind farm layout optimization (WFLO)
problem~\cite{MOSETTI1994105} is to locate wind turbines within a
predefined area to maximize the total power harvested from the wind
stream.  While physical considerations based on the diameter of the
blades prevent turbines being placed too close to each other, turbine
location decisions influence the total power production due to
interference effects (i.e., wakes) generated by upstream turbines.
These inter-turbine effects 
is best captured by Jensen's wake model~\cite{jensen1983note} when superimposed
using a sum-of-squares expression~\cite{Zhang2014}. However,
this wake modeling yields intractable optimization problems
that must be approximated using either quadratic or linear functions~\cite{turner2014new}.
An alternative formulation of WFLO relies on a less accurate wakes model, which uses linear superpositioning
to capture multiple wakes~\cite{donovan2005wind}.
The latter can be captured using quadratic
constraints and objective functions. However, exact optimization
approaches that have been proposed for the alternative WFLO formulation~\cite{Zhang2014,kuo2016wind,donovan2005wind,fagerfjall2010optimizing,archer2011wind,sorkhabi2018constrained}
represent the quadratic relations using linear models
that approximate wake dependencies, since until recent years solving linear programming problems
was computationally faster.

In this work, we introduce two novel quadratic WFLO models. The first
model represents WFLO with a quadratic constrained optimization
problem (QCOP).  The QCOP can be solved quickly using state-of-the-art solvers
that specialize in such problems, e.g., Gurobi~\cite{gurobi}.  Our
second model is based on a quadratic unconstrained binary optimization
(QUBO) representation of the WFLO problem.  Such a formulation enables the use of
specialized optimization hardware tailored to solve QUBOs. In this
paper, we use Fujitsu's Digital Annealer (DA), a hardware-based
QUBO solver, to demonstrate the strength of our approach. The main
contributions of this paper are:
\begin{enumerate} 
\item We propose a novel quadratic modeling framework for WFLO that captures wake 
effects directly through their linear superposition. 
%without the need for approximations and ad-hoc solutions,
\item We prove that WFLO is $\mathcal{NP}$-hard.
\item Through numerical experiments on two instantiations of our quadratic framework using Fujitsu's Digital Annealer and a commercial software-based quadratic solver (Gurobi), we demonstrate state-of-the-art performance in, respectively, quickly finding high-quality solutions and finding and proving optimal solutions over a longer run-time.
\end{enumerate}

@@

The remainder is structured as follows. Section~\ref{sec:related} presents our wind farm model and a literature review of exact and approximate
optimization solutions to the layout problem. The main contribution, namely the two quadratic models, is introduced in Section~\ref{sec:QUBO4WFLO}.
Section~\ref{sec:eval} demonstrates the usefulness of our approach via a thorough empirical evaluation of the proposed methods. 
Lastly, Section~\ref{sec:conclusion} provides concluding remarks and directions for future work.

@@

%\paragraph{Wind Farm as a Grid}
We consider a square-shaped wind farm that we model as an $n\times n$
grid with $n$ being the number of cells on each axis.  The cells have
an area of $c^2$ (in squared meters), i.e., if $c=100$ meters and
$n=20$, the farm area will be $2km\times 2km$.  In this work, we
assume that the terrain is flat and that we have no constraints
related to the terrain.  Furthermore, we do not account for the
influence of noise on the wind farm's surroundings. The wind farm
literature treats the more complex farm models that capture non-flat
terrain~\cite{song2015lazy,kuo2016wind}, multi-typed
turbines~\cite{feng2017design}, noise
effects~\cite{Zhang2014,sorkhabi2016impact,yin2014multi}, and complex
objectives (including turbine installation and maintenance
costs)~\cite{lackner2007analytical,sorkhabi2018constrained}.  In this work, we focus on the
novel quadratic approach for WFLO, and hence consider the most
abstract wind farm model. Our model can be extended, in the
future, to include additional wind farm settings.

@@

\paragraph{Effective Power Calculations in the Presence of Wakes} 
Let $\mathcal{D}$ be the set of wind regimes where every regime, $d\in \mathcal{D}$, is a combination of wind direction and free speed. 
The probability of wind regime 
$d \in \mathcal{D}$ is given by $p_d$ with $\sum_{d \in \mathcal{D}}^{} p_d = 1$. In the experimental section (Section~\ref{sec:eval}),
we provide one example of a probability distribution 
function over $36$ wind regimes (Figure~\ref{fig:prob_wind}). In that example, 
a regime $d=(10^\circ, 12 \mbox{ km/h})$ corresponds to a north-eastern wind blowing at a free velocity of $12\mbox{ km/h}$. The
probability of this regime is $0.008$.

@@

  Eq. (\ref{eq:ess}) is considered the most accurate analytical total power expression that accounts for multiple wakes~\cite{jensen1983note}. 
  However, the expression is not amenable to optimization since it is a squared root function of a cubic expression. 
  Hence, a simplified representation for the total expected energy that considers wakes is required.

@@

Note that $E_{LS}$ has the same parameters as $E_{SS}$ but differs in their relationship.
The LS model is less accurate compared to SS~\cite{Zhang2014}, yet as we will show in Section~\ref{sec:QUBO4WFLO} it 
results in a quadratic objective function in our WFLO formulations, which is a desired property from an optimization standpoint. \todo{[JCB: A bit strange to claim that a quadratic model is ``natural'' in the abstract but then argue that the model we adopt is an approximation but it is good because it is quadratic. Sounds like it is not natural and we choose it due to its optimization properties. Arik: I changed the arguments - please see the abstract and the beginning of the introduction.]}

@@

In what follows, we provide a literature review that covers both exact and approximate methods for WFLO. We subsequently relate our 
work to both approaches via our proposed quadratic optimization framework for WFLO.\footnote{We review 
only literature that considers the same wind farm model as the one presented at the beginning of this section. For a broader literature review see~\cite{Zhang2014}.}

\paragraph{Exact Optimization Methods} 
The WFLO problem can be solved  
to optimality, using
techniques from operations research (e.g., mixed-integer linear programming (MILP), quadratic programming (QP), and constraint programming (CP))~\cite{Zhang2014,turner2014new,donovan2005wind}.
These methods are \emph{exact}, in the sense that they guarantee that the returned solution is indeed
the best solution one can obtain for the given problem. Donovan was the first to solve WFLO using a
MILP approach based on the LS energy expression~\cite{donovan2005wind}. In~\cite{turner2014new}, the authors
approximated the SS energy expression using a quadratic and a linear approximation. The two resulting methods performed well
compared to existing approximate and earlier exact solutions.

The most recent attempt to solve the wind farm variation that we presented 
in this section is given in~\cite{Zhang2014}. The authors compared
previously existing MILP and QP models to their approach for directly optimizing the SS objective. To this end, they used constraint programming (CP), which 
is an optimization technique that does not make assumptions on the functional form of the objective.
However, applying CP turned out to be computationally intractable for the larger standard WFLO benchmarks.

In our work, we build upon the approach in~\cite{donovan2005wind}, yet instead of using a linear version of the LS objective, we 
model WFLO using a quadratic representation. The difference between our approach and the quadratic model
proposed in~\cite{turner2014new} is that our model is not an approximation of the SS objective, 
but an exact representation of the LS objective.

@@

\paragraph{Approximate Methods} In contrast to exact methods, approximation algorithms often provide   
quick and sufficiently accurate solutions. Several works solved 
WFLO using evolutionary (genetic) approaches~\cite{MOSETTI1994105,gonzalez2010optimization,grady2005placement}. 
These methods use two principles (cross-over and mutation)
to change existing solutions and improve them over time. Evolutionary methods do not guarantee termination 
(i.e., they may run forever without achieving a quality criteria), and when they do terminate (in practical cases) they do
not guarantee optimal solutions. Hence, one must often
introduce non-quality related stopping criteria, 
thus compromising on the quality of the obtained solution~\cite{davis1991handbook}.
Furthermore, there is no natural way of introducing constraints into genetic methods, which
leads to a potentially costly feasibility check for every solution found.

The first attempt to address these limitations 
was made using a local search approach for WFLO~\cite{ozturk2004heuristic}. 
The solution does not guarantee optimality, 
yet it circumvents the termination and feasibility checking issues of the evolutionary
methods. The local search approach attempts to apply either move, remove or add actions given an incumbent 
feasible solution. When a turbine is moved, removed or added, the new solution is evaluated. The procedure terminates 
when a stopping criteria is reached.
The main limitation of~\cite{ozturk2004heuristic} is that it cannot perform an action 
that leads to an infeasible state. Therefore, it quickly converges to a sub-optimal solution (local optimum) 
that can potentially be far
from the globally optimal solution. This problem was experimentally demonstrated in~\cite{rivas2009solving}.

To overcome this limitation, the authors of~\cite{rivas2009solving} used simulated annealing (SA), 
a neighborhood search method that allows visits to infeasible solutions. 
This creates a well-connected solution space that leads the algorithm to escape 
local optima. The SA method was shown to be superior to~\cite{ozturk2004heuristic} 
and to the genetic method proposed in~\cite{grady2005placement}. A drawback of the SA approach 
is its ad-hoc nature. The various components of the algorithm
 must be tailored and geared towards WFLO. This means that small adjustments of the WFLO setting
 (e.g., adding noise considerations), would require major changes to the SA solution. 
Additional methods and experimental comparisons between various approximate algorithms are available in~\cite{samorani2013wind}.

In our work, we also use an SA approach (similarly to~\cite{rivas2009solving}). However,
our methodology is generic since it is based on a model and
on the ability of specialized hardware to solve (non-ad-hoc) quadratic problems.

%\paragraph{Quadratic Programming as a Unified Framework for Exact and Approximate WFLO}
%\todo{send a clear message here. Perhaps move to next section}.

@@

In this work, we argue that 
formulating WFLO using quadratic programming enables us to employ 
advanced software solutions (e.g., Gurobi optimizer) and approximate hardware developments 
(e.g., Fujitsu's digital annealer that solves quadratic problems using principles from simulated annealing). 
A benchmark study compared state of the art exact approaches from~\cite{Zhang2014} to 
approximate methods and found that the two perform in a comparable fashion, without large differences between
the attained energy~\cite{yang2019simulated}. Exact methods tend to perform slightly better
in smaller WFLO benchmarks, while simulated annealing methods dominated for larger standard WFLO benchmarks. 
This leads to the conclusion that both types of methods (exact and approximate) should be considered. 
Therefore, in this work, we introduce a unified paradigm, namely quadratic programming, that enables
both exact and approximate solutions to WFLO. Thus, 
our approach `enjoys' the best of both worlds (the exact and the approximate).

In this section,
we formulate the WFLO problem
using quadratic programming. Specifically, we consider two
quadratic optimization problems that represent WFLO: a
quadratic constrained optimization (QCOP) problem and 
a quadratic unconstrained binary optimization (QUBO) problem. 
The former represents WFLO characteristics (proximity and number of turbines)
as hard constraints, while the latter considers these characteristics as soft constraints that are added to its objective function.

@@

Let $u_{id}$ and $u_{id\infty}$ 
be the wind speeds at turbine location $i \in \mathcal{N}$ for wind regime $d \in \mathcal{D}$
with and without interference from other turbines due to wake effects, respectively. 
Note that here we refer to $i$ as a potential turbine location and not the $i$th turbine as we did in Section~\ref{sec:related}.
Further, we denote by $\mathcal{U}_{id} \subseteq \mathcal{N}$ 
the set of upstream 
turbine locations for wind regime $d$
given that a turbine is placed at cell $i$ (i.e., turbines placed in these locations 
will be influenced by a turbine placed in $i$). Finally,
we denote by $u_{ijd}$ the wind speed at location $j$ due to a single wake from upstream turbine at location $i$ with $j \in \mathcal{U}_{id}$. 
The following function  
represents the total expected energy of a wind farm given placement decisions 
$x$: \begin{equation}
f(x) = \sum_{i \in \mathcal{N}}^{} \sum_{d \in \mathcal{D}}^{} p_d ( \frac{1}{3} \ u_{id, \infty}^3 x_i  - \sum_{j \in \mathcal{U}_{id}}^{} \frac{1}{3}(u_{id, \infty} ^3 - u_{ijd}^3)x_i x_j).   \label{ObjFunc}\\
%&s.t.:& \nonumber\\
%&\mbox{       }& \sum_{i \in \mathcal{N}}^{} x_i = m,\label{Cardinality}\\
%&\mbox{       }& x_i + x_j \leq 1,   \forall (i,j) \in \mathcal{E}, \label{Grid}\\
%&\mbox{       }& x_i \in \{0,1\},     \forall i \in \mathcal{N}.
\end{equation} The objective function corresponds to the linear superposition (LS)
expression (c.f., Eq.~\ref{eq:ls} in Section~\ref{sec:related}). We are now ready to provide the two quadratic programs to represent WFLO.

@@

To 
maximize $f(x)$ and satisfy the total number of turbines and proximity 
constraints we write
the following quadratic optimization problem (QCOP):
\begin{eqnarray} \label{QCOP}
&\max_{x_i}^{}& \sum_{i \in \mathcal{N}}^{} \sum_{d \in \mathcal{D}}^{} p_d ( \frac{1}{3} \ u_{id, \infty}^3 x_i  - \sum_{j \in \mathcal{U}_{id}}^{} \frac{1}{3}(u_{id, \infty} ^3 - u_{ijd}^3)x_i x_j)   \\
&s.t.:& \nonumber\\
&\mbox{       }& \sum_{i \in \mathcal{N}}^{} x_i = m,\label{Cardinality}\\
&\mbox{       }& x_i + x_j \leq 1,   \forall (i,j) \in \mathcal{E}, \label{Grid}\\
&\mbox{       }& x_i \in \{0,1\},     \forall i \in \mathcal{N}.
\end{eqnarray} Constraints~\ref{Cardinality}~and~\ref{Grid} ensure that exactly $m$ turbines are placed on the grid
and enforce proximity constraints between the turbines, respectively.

The QCOP presented as Problem~\ref{QCOP} can be solved using an exact optimization solver, e.g., Gurobi~\cite{gurobi}. 
So far, in~\cite{Zhang2014} and~\cite{donovan2005wind}, equivalent (linear) problems were formulated and solved using 
exact solvers. However, these solvers should only be used if the problem is computationally hard.
We are unaware of an existing complexity result for the LS variation of WFLO (our QCOP). In what follows, we provide
a theorem that proves the computational complexity of the LS-based WFLO QCOP. 
\begin{mythm}
	The computational complexity of QCOP is $\mathcal{NP}$-hard. 
\end{mythm}
\begin{myproof}
We prove by showing that a special case of the WFLO problem (Eq.~(\ref{QCOP}) is
$\mathcal{NP}$-hard. Suppose that $\mathcal{E} = \emptyset$, i.e., the grid is coarse grained enough to ignore proximity constraints. Then, the constraints in Eq.~(\ref{Grid})
are dropped and we get the heaviest k-subgraph problem (HSP), which was proven to be $\mathcal{NP}$-hard in~\cite{billionnet2005different}.
Since a special case of WFLO is  $\mathcal{NP}$-hard it implies that WFLO is at least $\mathcal{NP}$-hard. On the other hand,
WFLO is formulated using integer programming (Problem \ref{QCOP}), which means that it is at most $\mathcal{NP}$-hard. Therefore, we get that WFLO is $\mathcal{NP}$-hard.
	\end{myproof} To overcome the challenge associated with the computational hardness of WFLO,
we reformulate the QCOP into a QUBO problem, which allows us to solve WFLO using specialized optimization hardware.

@@

	\vspace{0.5em}
	\caption{Total Expected Power (kW/year) after 100 seconds as function of benchmark and method \todo{Change to 100 seconds and re-arrange our benchmarks first.} }\label{tab:results1}
\end{table}

\vspace{0.5em}
\caption{Total Expected Power (kW/year) after 3600 seconds as function of benchmark and method}\label{tab:results2}
\end{table}

@@

Future work: additional wind farm settings (noise, terrain...). Convert into qubo.
